ON USING THE DECISION TREES TO IDENTIFY THE LOCAL EXTREMA IN PARALLEL GLOBAL OPTIMIZATION ALGORITHM \footnote{The paper is recommended for publication by the Program Committee of the International Scientific Conference “Parallel Computational Technologies (PCT) 2023”}

### 2023-08-31 2023/vestnik.susu/example.tex
fix abstract

\title{ON USING THE DECISION TREES TO IDENTIFY THE LOCAL EXTREMA IN PARALLEL GLOBAL OPTIMIZATION ALGORITHMS
\footnote{The paper is recommended for publication by the Program Committee of the International Scientific Conference “Parallel Computational Technologies (PCT) 2023”}
}

@@

\begin{abstract}%
	In the present work, the solving of the multidimensional global optimization problems using decision tree to reveal the attractor regions of the local minima is considered. The objective function of the  problem is defined as a ``black box''.  We assume it only to satisfy Lipschitz condition with unknown constant.  Global search algorithm is applied for the search of global minimum in the problems of such type. In the present work, we propose a method for selecting the vicinity of local extrema of the objective  function based on analysis of accumulated search information. Conducting such an analysis using machine learning techniques allows making a decision to run a local  method, which can speed up the convergence of the algorithm.  This suggestion was confirmed by the results of numerical experiments demonstrating the speedup  when solving a series of test problems.

@@

\forcitationen{Pervyi~A.B., Vtoroi~V.G. The \LaTeXe{} Template for Paper Submission for the ``Computational Mathematics and Software Engineering'' Series}

@@

The main part of the paper has the following structure.  In Section \ref{SecPS}, the global optimization problem statement is considered, the basic suggestions  on the objective function are made. In Section \ref{SecA}, the description of the methods and approaches used is given. In particular, in Section \ref{SecGSA},  the main idea of global search algorithm is discussed, its computational rules are given. Section \ref{SecHG} contains the description of the local search method used. The method of constructing the approximation of the function using the decision trees is described in  Section \ref{SecDT}. The novel algorithm combining the global search and the local one on the base of using the decision  trees is presented in Section \ref{SecGSAL}. In the next section,  the features of the asynchronous parallelization scheme of the new algorithm are discussed. Section \ref{SecR} contains the results of experiments, carried out using a parallel computer system  with distributed memory, and a brief conclusion.

>>>>

\title{ON USING THE DECISION TREES TO IDENTIFY THE LOCAL EXTREMA IN PARALLEL GLOBAL OPTIMIZATION ALGORITHM
\footnote{The paper is recommended for publication by the Program Committee of the International Scientific Conference “Parallel Computational Technologies (PCT) 2023”}
}

@@

\begin{abstract}%
	In the present work, the solving of the multidimensional global optimization problems using decision tree to reveal the attractor regions of the local minima is considered. The objective function of the  problem is defined as a ``black box'', may be non-differentiable, multi-extremal and computational costly. We assume that the function satisfies the Lipschitz condition with a priory unknown constant.  Global search algorithm is applied for the search of global minimum in the problems of such type. It is well known that the solution complexity essentially depends on the presence of multiple local extrema. Within the framework of the global search algorithm, we propose a method for selecting the vicinity of local extrema of the objective  function based on analysis of accumulated search information. Conducting such an analysis using machine learning techniques allows making a decision to run a local  method, which can speed up the convergence of the algorithm.  This suggestion was confirmed by the results of numerical experiments demonstrating the speedup  when solving a series of test problems.

@@

\forcitationen{Barkalov~K.A., Lebedev~I.G., Silenko D.I. On using the decision trees to identify the local extrema in parallel global optimization algorithm}

@@

The main part of the paper has the following structure.  In Section \ref{SecPS}, the global optimization problem statement is considered, the basic suggestions  on the objective function are made. In Section \ref{SecA}, the description of the methods and approaches used is given. In particular, the main idea of global search algorithm is discussed and its computational rules are given. The section also contains the description of the local search method used and the method of constructing the approximation of the function using the decision trees. The novel algorithm combining the global search and the local one on the base of using the decision  trees is presented in Section \ref{SecGSAL}. In the next section,  the features of the asynchronous parallelization scheme of the new algorithm are discussed. Section \ref{SecR} contains the results of experiments, carried out using a parallel computer system  with distributed memory, and a brief conclusion.
